---
title: "4_Detailed_Report"
author: "Hyungjae Kim"
date: '2022-08-12'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, comment = NA, cache = T, fig.width=8, fig.height=4)

if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(ggplot2, dplyr, glmnet, car, tidyverse, gridExtra, ggrepel, plotly, skimr, data.table, xtable) 

set.seed(9)
```

```{r load}
dataset.raw <- fread("group1_hit_songs_dataset_raw1.csv")
dataset.1 <- fread("group1_hit_songs_dataset_ml_1.csv") 
dataset.5 <- fread("group1_hit_songs_dataset_ml_5.csv") 
```

## Dataset
### Raw Dataset
Removed `track_hist`, `url`, `artists_ids`, `track_id`, `chart_start`, `chart_end`, and `new_id`.
```{r 1 head}
head(dataset.1)
```

### Classified Dataset
Classified songs as hit songs by using a `hitness` threshold of 250, converted duration from milliseconds to minutes, and log scaled popularity and tempo. This dataset was used for all the models.
```{r 5 head}
head(dataset.5)
```

### Log Scaling Popularity
```{r log scale pop}
hist(dataset.1$popularity, main = "Artist Popularity Score", xlab = "popularity")
hist(dataset.5$log_popularity, main = "Log of Artist Popularity Score", xlab = "log(popularity)")
```

### Log Scaling Tempo
```{r log scale tempo}
hist(dataset.1$tempo, main = "Tempo", xlab = "tempo")
hist(dataset.5$log_tempo, main = "Log of Tempo", xlab = "log(tempo)")
```

### Number of Rows / Feature Vectors
```{r}
nrow(dataset.5)
```

### Summary of Raw Data (Continuous Hitness Values and Unscaled)
```{r summarize clean base}
summary(dataset.1)
```

### Summary of Classified and Cleaned Data (Factorized and Scaled)
```{r summarize ml base}
summary(dataset.5)
```

```{r data split}
n <- nrow(dataset.5)
dataset.5.train.index <- sample(n, n * 0.75)

dataset.5.train <- dataset.5[dataset.5.train.index, ]
dataset.5.test <- dataset.5.train <- dataset.5[-dataset.5.train.index, ]
```

## Linear Model (OLS)
```{r lm}
fit.lm.5 <- lm(hitness ~ ., dataset.5.train)
fit.lm.5.accuracy <- 1 - mean(abs(round(predict(fit.lm.5, dataset.5.test)) - dataset.5.test$hitness))
```

### Accuracy
```{r lm accuracy}
fit.lm.5.accuracy
```

### Summary (Coefficients, F-Statistics, P-Value)
```{r lm coef}
summary(fit.lm.5)
```

### Anova Table
```{r lm anova}
Anova(fit.lm.5)
```

### Plots (Residuals vs Fitted, Normal Q-Q, Scale-Location, Residuals vs Leverage)
```{r lm plot}
plot(fit.lm.5)
```

## Logistic Regression
```{r log reg}
fit.glm.5 <- glm(hitness ~ ., dataset.5.train, family = binomial(logit))

fit.glm.5.accuracy <- 1 - mean(abs(round(predict(fit.glm.5, dataset.5.test, type = "response")) - dataset.5.test$hitness))
```

### Accuracy
```{r log reg accuracy}
fit.glm.5.accuracy
```

### Summary (Coefficients, F-Statistics, P-Value)
```{r log reg coef}
summary(fit.glm.5)
```

### Anova Table
```{r log reg anova}
Anova(fit.glm.5)
```
### Plots (Residuals vs Fitted, Normal Q-Q, Scale-Location, Residuals vs Leverage)
```{r log reg plots}
plot(fit.glm.5)
```

```{r lasso}
x.5 <- data.matrix(dataset.5.train %>% select(-hitness))
y.5 <- dataset.5.train %>% select(hitness) %>% .$hitness

cv_model.5 <- cv.glmnet(x.5, y.5, alpha = 1)

fit.lasso.5 <- glmnet(x.5, y.5, alpha = 1, lambda = cv_model.5$lambda.1se)
fit.lasso.5.coef <- coef(fit.lasso.5)
fit.lasso.5.coef <- fit.lasso.5.coef[which(fit.lasso.5.coef != 0), ]
fit.lasso.5.names <- rownames(data.matrix(fit.lasso.5.coef))[-1]

predict.lasso.5 <- predict(fit.lasso.5, newx = data.matrix(dataset.5.test %>% select(-hitness)), type = "response")
fit.lasso.5.accuracy <- 1 - mean(abs(round(predict.lasso.5) - dataset.5.test$hitness))
```

## Logistic LASSO Regression
```{r log lasso reg, error = F}
fit.lasso.glm.5 <- glm(as.formula(paste("hitness ~", paste(all_of(fit.lasso.5.names), collapse = " + "))), dataset.5.train, family = binomial(logit))

predict.lasso.glm.5 <- predict(fit.lasso.glm.5, dataset.5.test %>% select(all_of(fit.lasso.5.names)), type = "response")
fit.lasso.glm.5.accuracy <- 1 - mean(abs(round(predict.lasso.glm.5) - dataset.5.test$hitness))
```

### Accuracy
```{r log lasso reg accuracy}
fit.lasso.glm.5.accuracy
```

### LASSO Log Selection (1SE)
```{r lasso plot}
plot(cv_model.5)
```

### Summary (Coefficients, F-Statistics, P-Value)
```{r log lasso reg coef}
summary(fit.lasso.glm.5)
```

### Anova Table
```{r log lasso reg anova}
Anova(fit.lasso.glm.5)
```

### Plots (Residuals vs Fitted, Normal Q-Q, Scale-Location, Residuals vs Leverage)
```{r log lasso reg plots}
plot(fit.lasso.glm.5)
```